Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-22 paper.qmd
Changed the gen_axes function

axes_obj <- gen_axes(
  data = scaled_gau_data,
  proj_data = projected_df,
  proj = projection,
  limits = 0.2,
  axis_pos_x = 0.15,
  axis_pos_y = -0.85)

@@

axes_obj <- gen_axes(
  data = scaled_gau_data,
  proj_data = projected_df,
  proj = projection,
  limits = 0.3,
  axis_pos_x = -0.7,
  axis_pos_y = -0.7)

@@

axes_obj <- gen_axes(
  data = scaled_gau_data,
  proj_data = projected_df,
  proj = projection,
  limits = 0.2,
  axis_pos_x = -0.55,
  axis_pos_y = -0.05)

@@

axes_obj <- gen_axes(
  data = scaled_gau_data,
  proj_data = projected_df,
  proj = projection,
  limits = 0.2,
  axis_pos_x = -0.37,
  axis_pos_y = 0.22)

@@

<!-- UMAP layout with author's suggested parameter choice-->
```{r}
#| label: published-pbmc
## Import data
training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
names(training_data_pbmc) <- paste0("p", 1:50)

@@

pbmc_model <- fit_highd_model(
  training_data = training_data_pbmc,
  emb_df = umap_pbmc_scaled,
  bin1 = num_bins_x_pbmc,
  r2 = r2_pbmc,
  q = 0.1,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "p"
)

@@

projected <- as.matrix(scaled_pbmc_data) %*% projection

@@

projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection

@@

axes_obj <- gen_axes(
  data = scaled_pbmc_data, 
  proj_data = projected_df, 
  proj = projection,
  limits = 1.5,
  axis_pos_x = -0.75,
  axis_pos_y = -0.75)

@@

pbmc_model <- fit_highd_model(
  training_data = training_data_pbmc,
  emb_df = umap_pbmc_scaled_best,
  bin1 = num_bins_x_pbmc,
  r2 = r2_pbmc,
  q = 0.1,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "p"
)

@@

## Compute error
error_df <- augment(
  df_bin_centroids = df_bin_centroids_pbmc,
  df_bin = df_bin_pbmc,
  training_data = training_data_pbmc,
  newdata = NULL,
  type_NLDR = "UMAP",
  col_start = "p")

@@

<!--bin1 = 22, bin2 = 17, b = 374, non_empty = 140-->
```{r}
#| label: mnist-model-proj
df_b_mnist <- df_bin_mnist |>
  dplyr::filter(hb_id %in% df_bin_centroids_mnist$hexID) |>
  dplyr::mutate(type = "model") ## Data with summarized mean

@@

## First projection
projection <- cbind(
    c(0.71503,0.00502,-0.04370,-0.01811,-0.01103,0.09877,0.06199,0.05396,0.07771,-0.01713),
    c(0.00587,0.71879,0.03485,0.08284,-0.07534,-0.05420,0.04315,-0.00547,-0.02995,0.01810)) * 1.3

projected <- as.matrix(scaled_mnist_data) %*% projection

@@

projected_model <- as.matrix(scaled_mnist_data_model) %*% projection

@@

axes_obj <- gen_axes(
  data = scaled_mnist_data, 
  proj_data = projected_df, 
  proj = projection,
  limits = 1.5,
  axis_pos_x = -0.75,
  axis_pos_y = -0.75)

@@

## Second projection
projection <- cbind(
    c(-0.33305,0.20223,0.06303,0.11124,0.13299,-0.30093,0.34082,0.15581,-0.08864,0.33573),
    c(0.53615,0.14969,-0.09393,-0.20053,0.07992,-0.34791,0.13057,-0.06814,-0.16737,0.03716)) * 2

projected <- as.matrix(scaled_mnist_data) %*% projection

@@

## Third projection
projection <- cbind(
    c(-0.12783,-0.50063,-0.02631,-0.00484,0.17733,0.37157,0.11310,0.09424,0.19844,0.19792),
    c(-0.42542,0.29570,0.31106,-0.30307,0.00614,0.16714,0.21779,-0.00025,0.07020,-0.00689)) * 1.5

projected <- as.matrix(scaled_mnist_data) %*% projection

>>>>

axes_obj <- gen_axes(  
  proj = projection,
  limits = 0.2,
  axis_pos_x = 0.15,
  axis_pos_y = -0.85,
  axis_labels = names(scaled_gau_data),
  threshold = 0.084)

@@

axes_obj <- gen_axes(  
  proj = projection,
  limits = 0.3,
  axis_pos_x = -0.7,
  axis_pos_y = -0.7,
  axis_labels = names(scaled_gau_data),
  threshold = 0.084)

@@

axes_obj <- gen_axes(  
  proj = projection,
  limits = 0.2,
  axis_pos_x = -0.55,
  axis_pos_y = -0.05,
  axis_labels = names(scaled_gau_data),
  threshold = 0.084)

@@

axes_obj <- gen_axes(  
  proj = projection,
  limits = 0.2,
  axis_pos_x = -0.37,
  axis_pos_y = 0.22,
  axis_labels = names(scaled_gau_data),
  threshold = 0.084)

@@

<!-- UMAP layout with author's suggested parameter choice-->
```{r}
#| label: published-pbmc
## Import data
training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
names(training_data_pbmc) <- paste0("pc", 1:50)

@@

pbmc_model <- fit_highd_model(
  training_data = training_data_pbmc,
  emb_df = umap_pbmc_scaled,
  bin1 = num_bins_x_pbmc,
  r2 = r2_pbmc,
  q = 0.1,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "pc"
)

@@

@@

@@

@@

pbmc_model <- fit_highd_model(
  training_data = training_data_pbmc,
  emb_df = umap_pbmc_scaled_best,
  bin1 = num_bins_x_pbmc,
  r2 = r2_pbmc,
  q = 0.1,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "pc"
)

@@

## Compute error
error_df <- augment(
  df_bin_centroids = df_bin_centroids_pbmc,
  df_bin = df_bin_pbmc,
  training_data = training_data_pbmc,
  newdata = NULL,
  type_NLDR = "UMAP",
  col_start = "pc")

@@

<!--bin1 = 22, bin2 = 17, b = 374, non_empty = 140-->
```{r}
#| label: mnist-model-proj

df_b_mnist <- df_bin_mnist |>
  dplyr::filter(hb_id %in% df_bin_centroids_mnist$hexID) |>
  dplyr::mutate(type = "model") ## Data with summarized mean

@@

## First projection
projection <- cbind(
    c(0.71503,0.00502,-0.04370,-0.01811,-0.01103,0.09877,0.06199,0.05396,0.07771,-0.01713),
    c(0.00587,0.71879,0.03485,0.08284,-0.07534,-0.05420,0.04315,-0.00547,-0.02995,0.01810))

projection_scaled <- projection * 1.3

projected <- as.matrix(scaled_mnist_data) %*% projection_scaled

@@

projected_model <- as.matrix(scaled_mnist_data_model) %*% projection_scaled

@@

axes_obj <- gen_axes(  
  proj = projection,
  limits = 1.5,
  axis_pos_x = -0.75,
  axis_pos_y = -0.75,
  axis_labels = names(scaled_mnist_data),
  threshold = 0.084)

@@

## Second projection
projection <- cbind(
    c(-0.33305,0.20223,0.06303,0.11124,0.13299,-0.30093,0.34082,0.15581,-0.08864,0.33573),
    c(0.53615,0.14969,-0.09393,-0.20053,0.07992,-0.34791,0.13057,-0.06814,-0.16737,0.03716))

projection_scaled <- projection * 2

projected <- as.matrix(scaled_mnist_data) %*% projection_scaled

@@

## Third projection
projection <- cbind(
    c(-0.12783,-0.50063,-0.02631,-0.00484,0.17733,0.37157,0.11310,0.09424,0.19844,0.19792),
    c(-0.42542,0.29570,0.31106,-0.30307,0.00614,0.16714,0.21779,-0.00025,0.07020,-0.00689))

projection_scaled <- projection * 1.5

projected <- as.matrix(scaled_mnist_data) %*% projection_scaled
